### 2005-11-10 doc/EFA/efa.tex
*** empty log message ***

The automaton $A$ is extended by a vector $X=(x_1,...,x_n)$ of $n$
finite integer variables $x_{k}$ and by a $|T|$-tuple of function
pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
(g_{\tau_{|T|}},a_{\tau_{|T|}})\big)$, where each pair
$(g_{\tau_k},a_{\tau_k})$ is associated to the corresponding
transition $\tau_k$ in $T$. $g_{\tau_k}$ is an evaluation of a
boolean formula (a guard formula) of $X$ i.e.
$g_{\tau_k}:{\mathbb{N}}^{n}\rightarrow \{0,1\}$ saying if the
corresponding transition $\tau_k$ is enabled or not. $a_{\tau_k}$
is an updating (action) function from ${\mathbb{N}}^{n}$ to
${\mathbb{N}}^{n}$ of the integer variables $X=(x_1,...,x_n)$.

With the above notations we will formally define an EFA as the
$8$-tuple $A_{e}=\langle Q, \Sigma, \delta_{e}, q_i, Q_m, X, G
,X_m \rangle$, where $X_m$ are the marked integer values of the
variables in $X$ and $\delta_{e}: Q \times \Sigma \times
{\mathbb{N}}^{n} \rightarrow Q\times {\mathbb{N}}^{n} $ is a new
extended version of the transition function $\delta:Q \times
\Sigma \rightarrow Q$. The extended transition function $\delta_e$
is defined by means of $\delta$ and the elements of $G$ as
\begin{eqnarray}
\delta_{e}(q,\sigma,X)=\left\{
\begin{array}{ll}
(\delta(q,\sigma),a_{q \sigma}(X)\big)\quad\quad g_{q \sigma}(X)=1\\
\textrm{undefined} \quad\quad\quad\quad\quad g_{q \sigma}(X)=0
\end{array}\right.
\end{eqnarray}
If $(q,\sigma)\notin T$ then $g_{q\sigma}$ is undefined and
therefore $\delta_{e}(q,\sigma,X)$ is undefined. Notice that for
the set of transitions $T_e$ in the extended automaton $A_e$ we
have $T_e=\{(q,\sigma)\in Q \times \Sigma: g_{q\sigma}(X)=1
\textrm{ for some } X\}\subset \{(q,\sigma)\in Q \times \Sigma:
g_{q\sigma} \textrm{ is defined } \}\ =T$.

\subsection{Synchronous composition of EFA's}
The synchronous composition of two EFA's $A=\langle Q^A, \Sigma^A,
\delta^A, q^A_i, Q^A_m, X^{A}, G^A, X^A_m \rangle$ and $B=\langle
Q^B, \Sigma^B, \delta^B, q^B_i, Q^B_m, X^{B}, G^B, X^B_m \rangle$
exists if and only if the automata are consistent. The consistency
condition requires that the action functions triggered by a common
enabled event $\sigma\in \Sigma_{A}\cap \Sigma_{B}$ in the two
automata update common variables to the same values, i.e. if the
$i^{ith}$ variable in $X^{A}$ and the $j^{ith}$ variable in
$X^{B}$ are the same variable and $g_{q^A \sigma}(X^{A})\wedge
g_{q^B \sigma}(X^{B})=1$ where $(q^A,q^B)\in Q^A\times Q^B$ then
$(a_{q^A \sigma}(X^{A}))_i = (a_{q^B \sigma}(X^{B}))_j$.

The consistency condition can be verified by inspection. However,
if the vectors $X^{A}$ or $X^{B}$ disable the corresponding guard
formula ($g_{q^A \sigma}(X^{A})\wedge g_{q^B \sigma}(X^{B})$) it
may happen that $a_{q^A \sigma}(X^{A}))_i \neq (a_{q^B
\sigma}(X^{B}))_j$ even though the two automata are consistent.
Therefore it may be a time-consuming task to verify the
consistency condition by inspection. In section
\ref{Implementation} we will present algorithms that includes a
consistency check of the synchronous
composition of two EFA's. \\

@@

 Let $Z^{AB}$ denote the common variables of $X^{A}$ and $X^{B}$, and let $Z^A$
and $Z^B$ be the distinct variables of $X^{A}$ and $X^{B}$,
respectively. We define the integer variables $Z$ of the
synchronous product of $A$ and $B$ as the vector $Z=(Z^A, Z^{AB},
Z^B)$. The guard functions $g_{q \sigma}(Z)$ are defined as
\begin{eqnarray}\label{guard_sync_1}
g_{q \sigma}(Z)=\left\{
\begin{array}{ll}
g_{q^A\sigma}(X^{A})\wedge g_{q^B\sigma}(X^{B}) \textrm{ if }\sigma\in \Sigma^A \cap \Sigma^B\\
g_{q^A\sigma}(X^{A}) \quad\quad \textrm{ if } \sigma\in \Sigma^A \setminus \Sigma^B\\
g_{q^B\sigma}(X^{B}) \quad\quad \textrm{ if }
\sigma\in\Sigma^B\setminus \Sigma^A
\end{array}\right.
\end{eqnarray}
If any one of the guards functions in the above definition are
undefined the guard function for the product $g_{q \sigma}$ is
also undefined. To simplify the notation further, we use "dummy"
guard functions in the following way. If $\sigma\in \Sigma^A
\setminus \Sigma^B$ we introduce fictional guards functions
$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. With this convention we can
write equation (\ref{guard_sync_1}) as
\begin{eqnarray}
g_{q \sigma}(Z)= g_{q^A\sigma}(X^{A})\wedge g_{q^B\sigma}(X^{B}).
\end{eqnarray}
To define the action functions for $A||B$ we look at the
restrictions of the updating functions $a^A_\delta$ and
$a^B_\delta$ to the sets $Z^A$ and $Z^B$. The updating functions
are defined as (if $X^{A}$ and $X^{B}$ have no common variables
then $Z^A=X^{A}$ and $Z^B=X^{B}$)
\begin{eqnarray}
a_{q \sigma}(Z)=\left\{
\begin{array}{ll}
 \big(a^A_{q^A \sigma}(X^{A}), a^B_{q^B \sigma}(Z^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
\big(a^A_{q^A \sigma}(X^{A}), Z^{B}\big)\quad\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
\big(Z^{A}, a^B_{q^B \sigma}(X^{B})\big)\quad\quad\quad\sigma\in
\Sigma^B \setminus \Sigma^A
\end{array}\right.
\end{eqnarray}
As in the definition of the guard functions the notion of being
undefined is inherited from the action functions of $A$ and $B$.
Similar to the "dummy" guard functions we introduce fictional
action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
introduce fictional action functions $a_{q^B\sigma}(X^B)=X^B$ for
$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
$a_{q^A\sigma}(X^A)=X^A$. If $X^{A}$ and $X^{B}$ have no common
variables then the formula for the action functions of the product
simplifies to
\begin{eqnarray}
a_{q \sigma}(Z)=
 \big(a^A_{q^A \sigma}(X^{A}), a^B_{q^B \sigma}(X^{B})\big).
\end{eqnarray}

 The transition function
$\delta(q,\sigma,Z)$ is defined from $\delta^A$ and $\delta^B$ as
\begin{eqnarray}
\delta(q,\sigma,Z)=\left\{
\begin{array}{ll}
\big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(Z)\big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(Z)=1\\
\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(Z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(Z)=1\\
\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(Z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(Z)=1\\
\textrm{
undefined}\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
g_{q\sigma}(Z)=0
\end{array}\right.
\end{eqnarray}

>>>>

The automaton $A$ is extended by a set $X=X_1\times ... \times
X_n$, where $X_k$, $k=1,...,n$, are a finite nonempty subsets of
$\mathbb{Z}$. The set $X$ is the domain of definition of an
$n$-tuple of integer variables $x=(x_1,...,x_n)$ with initial
value $x_0=(x_{10},...,x_{n0})$. We also add to $A$ an $|T|$-tuple
of function pairs $G=\big((g_{\tau_1},a_{\tau_1}),...,
(g_{\tau_{|T|}},a_{\tau_{|T|}})\big)$, where each pair
$(g_{\tau_k},a_{\tau_k})$ is associated to the corresponding
transition $\tau_k$ in $T$. $g_{\tau_k}$ is an evaluation of a
boolean formula (a guard formula) over $X$ i.e. $g_{\tau_k}:X
\rightarrow \{0,1\}$ saying if the corresponding transition
$\tau_k$ is enabled or not. $a_{\tau_k}$ is an updating (action)
function from $X$ to $X$ of the integer variables
$x=(x_1,...,x_n)$.

With the above notations we will formally define an EFA as the
$9$-tuple $A_{e}=\langle Q, \Sigma, \delta_{e}, q_i, Q_m, X, G
,x_0, X_m \rangle$, where $X_m$ are the marked integer values of
the variables $x$ and $\delta_{e}: Q \times \Sigma \times X
\rightarrow Q\times X $ is a new extended version of the
transition function $\delta:Q \times \Sigma \rightarrow Q$. The
extended transition function $\delta_e$ is defined by means of
$\delta$ and the elements of $G$ as
\begin{eqnarray}
\delta_{e}(q,\sigma,x)=\left\{
\begin{array}{ll}
(\delta(q,\sigma),a_{q \sigma}(x)\big)\quad\quad g_{q \sigma}(x)=1\\
\textrm{undefined} \quad\quad\quad\quad\quad g_{q \sigma}(x)=0
\end{array}\right.
\end{eqnarray}
If $(q,\sigma)\notin T$ then $g_{q\sigma}$ is undefined and
therefore $\delta_{e}(q,\sigma,x)$ is undefined. Notice that for
the set of transitions $T_e$ in the extended automaton $A_e$ we
have $T_e=\{(q,\sigma)\in Q \times \Sigma: g_{q\sigma}(x)=1
\textrm{ for some } x\in X\}\subset \{(q,\sigma)\in Q \times
\Sigma: g_{q\sigma} \textrm{ is defined } \}\ =T$.

\subsection{Synchronous composition of EFA's}
The synchronous composition of two EFA's $A=\langle Q^A, \Sigma^A,
\delta^A, q^A_i, Q^A_m, X^{A}, G^A, X^A_m \rangle$ and $B=\langle
Q^B, \Sigma^B, \delta^B, q^B_i, Q^B_m, X^{B}, G^B, X^B_m \rangle$
exists if and only if the automata are consistent. The consistency
condition requires that the action functions triggered by a common
enabled event $\sigma\in \Sigma_{A}\cap \Sigma_{B}$ in the two
automata update common variables to the same values, i.e. if the
$i^{ith}$ variable in $x^{A}$ and the $j^{ith}$ variable in
$x^{B}$ are the same variable and $g_{q^A \sigma}(x^{A})\wedge
g_{q^B \sigma}(x^{B})=1$ where $(q^A,q^B)\in Q^A\times Q^B$ then
$(a_{q^A \sigma}(x^{A}))_i = (a_{q^B \sigma}(x^{B}))_j$.

The consistency condition can be verified by inspection. However,
if the vectors $x^{A}$ or $x^{B}$ disable the corresponding guard
formula ($g_{q^A \sigma}(x^{A})\wedge g_{q^B \sigma}(x^{B})$) it
may happen that $a_{q^A \sigma}(x^{A}))_i \neq (a_{q^B
\sigma}(x^{B}))_j$ even though the two automata are consistent.
Therefore it may be a time-consuming task to verify the
consistency condition by inspection. In section
\ref{Implementation} we will present algorithms that includes a
consistency check of the synchronous
composition of two EFA's. \\

@@

 Let $z^{AB}$ denote the common variables of $x^{A}$ and $x^{B}$, and let $z^A$
and $z^B$ be the distinct variables of $x^{A}$ and $x^{B}$,
respectively. We define the integer variables $z$ of the
synchronous product of $A$ and $B$ as the vector $z=(z^A, z^{AB},
z^B)$. The guard functions $g_{q \sigma}(z)$ are defined as
\begin{eqnarray}\label{guard_sync_1}
g_{q \sigma}(z)=\left\{
\begin{array}{ll}
g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}) \textrm{ if }\sigma\in \Sigma^A \cap \Sigma^B\\
g_{q^A\sigma}(x^{A}) \quad\quad \textrm{ if } \sigma\in \Sigma^A \setminus \Sigma^B\\
g_{q^B\sigma}(x^{B}) \quad\quad \textrm{ if }
\sigma\in\Sigma^B\setminus \Sigma^A
\end{array}\right.
\end{eqnarray}
If any one of the guards functions in the above definition are
undefined the guard function for the product $g_{q \sigma}$ is
also undefined. To simplify the notation further, we use "dummy"
guard functions in the following way. If $\sigma\in \Sigma^A
\setminus \Sigma^B$ we introduce fictional guards functions
$g_{q^B\sigma}(.)=1$ for $B$ and if $\sigma\in\Sigma^B\setminus
\Sigma^A$ we set $g_{q^A\sigma}(.)=1$. With this convention we can
write equation (\ref{guard_sync_1}) as
\begin{eqnarray}
g_{q \sigma}(z)= g_{q^A\sigma}(x^{A})\wedge g_{q^B\sigma}(x^{B}).
\end{eqnarray}
To define the action functions for $A||B$ we look at the
restrictions of the updating functions $a^A_\delta$ and
$a^B_\delta$ to the sets $Z^A$ and $Z^B$. The updating functions
are defined as (if $x^{A}$ and $x^{B}$ have no common variables
then $z=(x^{A},x^{B})$)
\begin{eqnarray}\label{action_sync_1}
a_{q \sigma}(z)=\left\{
\begin{array}{ll}
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(z^{B})\big)\quad\quad\sigma\in \Sigma^A \cap \Sigma^B\\
\big(a^A_{q^A \sigma}(x^{A}), z^{B}\big)\quad\quad\quad\sigma\in \Sigma^A \setminus \Sigma^B\\
\big(z^{A}, a^B_{q^B \sigma}(x^{B})\big)\quad\quad\quad\sigma\in
\Sigma^B \setminus \Sigma^A
\end{array}\right.
\end{eqnarray}
As in the definition of the guard functions the notion of being
undefined is inherited from the action functions of $A$ and $B$.
Similar to the "dummy" guard functions we introduce fictional
action functions. If $\sigma\in \Sigma^A \setminus \Sigma^B$ we
introduce fictional action functions $a_{q^B\sigma}(x^B)=x^B$ for
$B$ and if $\sigma\in\Sigma^B\setminus \Sigma^A$ we set
$a_{q^A\sigma}(x^A)=x^A$. If $x^{A}$ and $x^{B}$ have no common
variables then equation (\ref{action_sync_1}), that defines the
action functions of the product, simplifies to
\begin{eqnarray}
a_{q \sigma}(z)=
 \big(a^A_{q^A \sigma}(x^{A}), a^B_{q^B \sigma}(x^{B})\big).
\end{eqnarray}

 The transition function
$\delta(q,\sigma,z)$ is defined from $\delta^A$ and $\delta^B$ as
\begin{eqnarray}
\delta(q,\sigma,z)=\left\{
\begin{array}{ll}
\big(\delta^A(q^A,\sigma),\delta^B(q^B,\sigma),a_{q \sigma}(z)\big) \quad\quad \sigma\in \Sigma^A \cap \Sigma^B \quad g_{q \sigma}(z)=1\\
\big(\delta^A(q^A,\sigma),q^B,a_{q \sigma}(z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^A \setminus \Sigma^B \quad g_{q \sigma}(z)=1\\
\big(q^A,\delta^B(q^B,\sigma),a_{q \sigma}(z)\big) \quad\quad\quad\quad\quad \sigma\in \Sigma^B \setminus \Sigma^A \quad g_{q \sigma}(z)=1\\
\textrm{
undefined}\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
g_{q\sigma}(z)=0
\end{array}\right.
\end{eqnarray}
